---
title: "R Notebook"
output: html_notebook
---




```{r}
setwd("C:/Users/brian/OneDrive/3 Data Science/Data Science Project")
```
#Importing the packages needed 
```{r}

Packages <- c('car', "caret", 'corrplot', 'datasets', 'data.table', 'dplyr', 'ggplot2', 'graphics', 'knitr', 'maps', 'mice','plyr', 'plotly', 'reshape', 'stats', 'usmap', 'VIM', 'xlsx', 'xlsxjars')
lapply(Packages, library, character.only = TRUE)

#library(rjava)
```

#Loading the external data

```{r}
#1
county_facts <- read.csv("county_facts.csv")
#2
pres16res <- read.csv("pres16results.csv")
#3
pe16rwvd9 <- read.csv("pres.elect16.results.wide.votes.dec9.csv")
#4
primres <- read.csv("primary_results.csv")
#5
exitpolls <- read.csv("exitpollsdatacnn04022017.csv")
#6
votesRA = read.csv("Voteperstate.csv")
#7
electoral = read.csv("2016_electoral_votes.csv")
#8
censusRA = read.csv("censusraceagevote.csv")
#9
#elec2012 <- read.csv("2012_pres_results.txt")

data(county.fips, state.fips)
states <- map_data("state")
counties <- map_data("county")
```

#Testing each data set for missingness
```{r}

#County facts
aggr1_plotfacts <- aggr(county_facts, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('County Facts Histogram of Missing Data','Pattern'))

#presidential election 2016 results
aggr1pres <- aggr(pres16res, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Presidential Results Histogram of Missing Data','Pattern'))

#wide results
aggr1wide <- aggr(pe16rwvd9, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Wide Presidential Results Histogram of Missing Data','Pattern'))

#primary dataset
aggr1prim <- aggr(primres, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Primary Result Histogram of Missing Data','Pattern'))

#Exit polls
aggrexit <- aggr(exitpolls, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Exit Polls Histogram of Missing Data','Pattern'))
```

#CLEANING the data/ Datframe manipulation
#1: County Facts
```{r}
#county_facts=county_facts[, !(colnames(county_facts) %in% c("AGE135214", "LFE305213", "MAN450207", "WTN220207", "LND110210"))]

#Renaming the columns for easier reading
setnames(county_facts, old=c("PST045214", "state_abbreviation", "RHI125214","RHI225214", "RHI325214", "RHI425214", "RHI525214", "RHI625214", "RHI725214", 'RHI825214'), new=c("Population", "sa", "White", "Black", "Native American", "Asian", "Pacific Islander", "2+ Races", "Latino", "White/Not Latino"))

#Creating a states Dataframe, while removing the state facts from the county_facts data set

stfips = data.frame(fips=1000*state.fips$fips)
#Adding non-continental states to data frame: Alaska 2000, Hawaii 15000
stfips[64,] <- 2000
stfips[65,] <- 15000

#Creating the states dataframe
statesfacts = county_facts[(county_facts$fips %in% stfips$fips),]
statesfacts$state_abbreviation <- NULL
statesfacts=statesfacts[-c(3)]

#removing state facts
county_facts = county_facts[!(county_facts$fips %in% stfips$fips),]
```
#2: Presidential Results 16
```{r}
#Separating the state and the county results
nodata <- data.frame(pres16res[18352:18475, 1:2])

NoResData <- unique(pres16res[is.na(pres16res[c(3:9)]),])

 
pres16resST = pres16res[is.na(pres16res$county),]
pres16resCT = pres16res[!is.na(pres16res$county),]
```

#3: Pres.election 16 results Dec.9 Wide Votes
```{r}
regrpesubset = subset(pe16rwvd9)
#regrpesubset[is.na(regrpesubset[c(8:39)])] <- 0
regrpesubset =regrpesubset[with(regrpesubset, order(regrpesubset$fips)), ]
rownames(regrpesubset) <- 1:nrow(regrpesubset)

statesvotes<-subset(pe16rwvd9,  is.na(pe16rwvd9$county)) 

#Replacing original county data without state data
pe16rwvd9 = pe16rwvd9[!(pe16rwvd9$county %in% statesvotes$county),]
pe16rwvd9 =pe16rwvd9[with(pe16rwvd9, order(pe16rwvd9$fips)), ]
rownames(pe16rwvd9) <- 1:nrow(pe16rwvd9)

setnames(pe16rwvd9, old = c(8:13, 15:19, 21:39), new = c("other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other"))

#Making the other candidate & votes column
pe16rwvd9 <- unique(melt(pe16rwvd9, id=c(1:7,14,20), na.rm = FALSE))
pe16rwvd9 <- pe16rwvd9[order(pe16rwvd9$fips) , ]

#Changing the names of these 
setnames(pe16rwvd9, old = c("value", "variable"), new= c("Votes", "Candidates"))

#Removing excess blank/redundant columns
setnames(statesvotes, old = c(8:13, 15:19, 21:39), new = c("other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other", "other"))

#Making the other candidate & votes column
statesvotes <- unique(melt(statesvotes, id=c(1:7,14,20), na.rm = FALSE))
statesvotes <- statesvotes[order(statesvotes$state.name) , ]

#Removing excess/blank columns
statesvotes <- statesvotes[c(-(53:54)), c(-(1:4))]

#Changing the names of these 
setnames(statesvotes, old = c("value", "variable"), new= c("Votes", "Candidates"))

#Reordering columns and row indices
statesvotes <- statesvotes[c(3, 2, 1, 5, 4, 6, 7 )]
rownames(statesvotes) <- 1:nrow(statesvotes)
```

#5: Exit Polls
```{r}
exitpolls <- unique(exitpolls[-c(1, 10)])
```

#6: Votes by Race and Age
```{r}
#Cleaning 
votesRA  <- subset(votesRA, select = -c(2, 21:22, 25))
setnames(votesRA, 
         old = c("ï..State.or..district", "Hillary.Clinton..Democratic.."  , "Hillary.Clinton..Democratic...1" ,       "Hillary.Clinton..Democratic.Electoral..votes",  "Donald.Trump..Republican..","Donald.Trump..Republican...1" ,"Donald.Trump..Republican.Electoral..votes"  ,  "Gary.Johnson..Libertarian.." ,"Gary.Johnson..Libertarian...1", "Gary.Johnson..Libertarian.Electoral..votes","Jill.Stein..Green..","Jill.Stein..Green...1","Jill.Stein..Green.Electoral..votes","Evan.McMullin..Independent.." ,"Evan.McMullin..Independent...1" ,"Evan.McMullin..Independent.Electoral..votes","Others.." ,"Others...1" ,"Others.Electoral..votes","Total.." ,"Total.State"), 
         new = c("State/District", "HC Vote #", "HC Vote Perc","HC E.Votes", "DT Vote #", "DT Vote Perc", "DT E.Votes", "GJ Vote #", "GJ Vote Perc","GJ E.Votes", "JS Vote #", "JS Vote Perc", "JS E.Votes", "EM Vote #", "EM Vote Perc","EM E.Votes", "Others Vote #", "Others Vote Perc", "Other E.Votes", "Total Vote #", "State"))  
```
#7: Electoral Votes data
```{r}
setnames(censusRA, old=c("ï..Race"), new = c('CensusRace'))
setnames(statesfacts, old=c("area_name"), new =c("state"))

electoral = merge(electoral, statesfacts[c(1:2)], by= "state")
```

#Checking for missing values after cleaning
```{r}
#County facts
aggr1_plotfacts <- aggr(county_facts, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('County Facts Histogram of Missing Data','Pattern'))

#presidential election 2016 results
aggr1pres <- aggr(pres16res, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Presidential Results Histogram of Missing Data','Pattern'))

#wide results
aggr1meltwide <- aggr(pe16rwvd9, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Wide Presidential Results Histogram of Missing Data','Pattern'))

#primary dataset
aggr1prim <- aggr(primres, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Primary Result Histogram of Missing Data','Pattern'))

#Exit polls
aggrexit <- aggr(exitpolls, col = c('Blue','Purple'), numbers = TRUE, sortVars = TRUE, cex.axis = .8, gap = 3, ylab = c('Exit Polls Histogram of Missing Data','Pattern'))


```

#Making other dataframe from the originals
#1: County Facts Race Population Data
```{r}
USRacePop = county_facts[c(1), c(2, 4, 12:19)]

#Eliminating the US total facts row
race = county_facts[-1,c(1:4, 12:19)]

#Converting percentages to decimal 
race[c(5:12)] <-replace(race[c(5:12)], values =0.01*race[c(5:12)])
race[c(5:12)] <-replace(race[c(5:12)], values =race$Population *race[c(5:12)])
```

#5: Exit polls data
```{r}
eprace <- exitpolls[exitpolls$questions=="Race" | exitpolls$questions=="Race and gender" | exitpolls$questions=="Education and race" | exitpolls$questions=="Age by race",]

eprace[c(4:7)] <-replace(eprace[c(4:7)], values =0.01*eprace[c(4:7)])

eprace =eprace[with(eprace, order(eprace$questions)), ]
rownames(eprace) <- 1:nrow(eprace)


```

#Barplots of Trump & Clinton Support by race
```{r}
racetitle <- c("White", "Black", "Latino", "Asian", "Other", 'Non-White')

racecounts <- data.frame(racetitle, h= (eprace[eprace$questions=="Race" & eprace$state=="nation", c(4)]), t= (eprace[eprace$questions=="Race" & eprace$state=="nation", c(5)]))

p <- plot_ly(racecounts, x = ~racetitle, y = ~h , type = 'bar', name = 'Hillary Support', marker = list(color = 'blue')) %>%
  add_trace(y = ~t, name = 'Trump Support',marker = list(color = 'red')) %>%
  layout(yaxis = list(title = 'Percentage of Votes'), xaxis = list(title = 'Racial/Ethnic Groups') , barmode = 'stack') 
p

```
#Barplots of Trump & Clinton Support by race & gender
```{r}
racegentitle <- c('White men', 'White women', 'Black men', 'Black women', 'Latino men', 'Latino women', 'Others')

racegencounts <- data.frame(racegentitle, h= (eprace[eprace$questions=="Race and gender" & eprace$state=="nation", c(4)]), t= (eprace[eprace$questions=="Race and gender" & eprace$state=="nation", c(5)]))

p <- plot_ly(racegencounts, x = ~racegentitle, y = ~h , type = 'bar', name = 'Hillary Support', marker = list(color = 'blue')) %>%
  add_trace(y = ~t, name = 'Trump Support',marker = list(color = 'red')) %>%
  layout(yaxis = list(title = 'Percentage of Votes by Gender & Race'), xaxis = list(title = 'Racial/Ethnic Groups') , barmode = 'stack') 
p

```
#Barplots of Trump & Clinton Support by race & age
```{r}
raceagetitle <- c('Whites 18-29', 'Latinos 30-44', 'Latinos 45-64', 'Latinos 65 and older', 'All others', 'Whites 30-44', 'Whites 45-64', 'Whites 65 and older', 'Blacks 18-29', 'Blacks 30-44', 'Blacks 45-64',	 'Blacks 65 and older', '13	Latinos 18-29')

raceagecounts <- data.frame(raceagetitle, h= (eprace[eprace$questions=="Age by race" & eprace$state=="nation", c(4)]), t= (eprace[eprace$questions=="Age by race" & eprace$state=="nation", c(5)]))

raceagecounts = raceagecounts[c(5, 13, 2, 3, 4, 1, 6, 7, 8, 9, 10, 11, 12), ]

p <- plot_ly(raceagecounts, x = ~raceagetitle, y = ~h , type = 'bar', name = 'Hillary Support', marker = list(color = 'blue')) %>%
  add_trace(y = ~t, name = 'Trump Support',marker = list(color = 'red')) %>%
  layout(yaxis = list(title = 'Percentage of Votes by Race & Age'), xaxis = list(title = "", tickangle = 45),  barmode = 'stack', margin = list(b = 200)) 
p

```
#Barplots of Trump & Clinton Support by race & education
```{r}
eduracetitle <- c('White college graduates', 'Whites no degree', 'Non-whites college grads', 'Non-whites no degree')

eduracecounts <- data.frame(eduracetitle, h= (eprace[eprace$questions=="Education and race" & eprace$state=="nation", c(4)]), t= (eprace[eprace$questions=="Education and race" & eprace$state=="nation", c(5)]))


p <- plot_ly(eduracecounts, x = ~eduracetitle, y = ~h , type = 'bar', name = 'Hillary Support', marker = list(color = 'blue')) %>%
  add_trace(y = ~t, name = 'Trump Support',marker = list(color = 'red')) %>%
  layout(yaxis = list(title = 'Percentage of Votes by Race and Education'), xaxis = list(title = "", tickangle = 45),  barmode = 'stack', margin = list(b = 200)) 
p

```

#Barplot of counties won between HC and DT
```{r}

ggplot(data = pe16rwvd9, aes(x = lead, fill = lead)) + geom_bar() + xlab("") + 
    ylab("") + scale_fill_manual(name = "Main Candidate", values = rainbow(2))
```

#Barplot of Total counties per state
```{r}
bar1 <- ggplot(data=race, aes(x=sa))
bar1 + geom_bar() +   xlab("Counties") +  ylab("Number of Counties") + ggtitle("Counties per State") 
```

#Forming the main candidate dataframe which consists of total votes, votes per candidate, as well as percentage of votes and the racial population percentages per county
```{r}
hcperc = pe16rwvd9$Hillary.Clinton/pe16rwvd9$total_votes
dtperc = pe16rwvd9$Donald.Trump/pe16rwvd9$total_votes

maincands=subset(pe16rwvd9[-1,], select = c(fips, st, county, total_votes, lead, Hillary.Clinton, Donald.Trump))

maincands=cbind(maincands, hcperc[-1], dtperc[-1])

maincands= maincands[c(1, 2,3, 4, 5, 7, 6, 8)]

wperc = race$White/race$Population
bperc = race$Black/race$Population
naperc = race$`Native American`/race$Population
aperc = race$Asian/race$Population
piperc = race$`Pacific Islander`/race$Population
biperc = race$`2+ Races`/race$Population
lperc = race$Latino/race$Population
eperc = race$`White/Not Latino`/race$Population

raceperc = data.frame(cbind(race['fips'],wperc, bperc, naperc, aperc, piperc, biperc, lperc, eperc))

setnames(raceperc, old=c('wperc', 'bperc', 'naperc', 'aperc', 'piperc', 'biperc', 'lperc', 'eperc'), new=c("White Perc", "Black Perc", "Native American Perc", "Asian Perc", "Pacific Islander Perc", "2+ Races Perc", "Latino", "White/Not Latino Perc"))


rcs =merge(maincands, raceperc, by = "fips", all = FALSE)[-1]
rcs =rcs[with(rcs, order(rcs$st)), ]
rownames(rcs) <- 1:nrow(rcs)
```


#Forming the state population by race dataframe
```{r}
staterace = data.frame(statesfacts[1:3], statesfacts[11:18]*0.01)

wstate = staterace['White']*staterace['Population']
bstate = staterace['Black']*staterace['Population']
nastate = staterace['Native.American']*staterace['Population']
astate = staterace['Asian']*staterace['Population']
pistate = staterace['Pacific.Islander']*staterace['Population']
mustate = staterace['X2..Races']*staterace['Population']
lstate = staterace['Latino']*staterace['Population']
estate = staterace['White.Not.Latino']*staterace['Population']

staterace = data.frame(staterace[1:3], wstate, bstate,nastate,astate,pistate,mustate,lstate,estate)
```

```{r}
othershil = (staterace[6]+ staterace[8]+ staterace[9])*0.56
hilcount = data.frame(staterace[2], staterace[4]*0.37, staterace[5]*0.89,staterace[10]*0.66, staterace[7]*0.65, othershil)
setnames(hilcount, old=c("area_name"), new=c("state.name"))

hilrcs = merge(statesvotes, hilcount, by= "state.name", all = FALSE)[-1]

hilrcs = cbind(hilrcs, StateName=statesvotes$state.name)
hilrcs = hilrcs[c(12, 1:11)]
hilrcs = hilrcs[with(hilrcs, order(hilrcs$StateName)), ]
rownames(hilrcs) <- 1:nrow(hilrcs)

```

```{r}
colnames(hilrcs)
```





#Logistic Regression on the candidates
```{r}
rcsonlyrace = na.omit(rcs[c(3:15)])
sapply(rcsonlyrace,function(x) sum(is.na(x)))
sapply(rcsonlyrace, function(x) length(unique(x)))
library(Amelia)
missmap(rcsonlyrace, main = "Missing values vs observed")
is.factor(rcsonlyrace$lead)
contrasts(rcsonlyrace$lead)
```

```{r}
train <- rcsonlyrace[1:2789,]
test <- rcsonlyrace[2799:3110,]

model <- glm(lead ~.,family=binomial(link='logit'), data=train)
summary(model)

```
```{r}
p<-predict(model, test, type = "response")

p_class <- ifelse(p<1, "D", "H")


table(p_class, test$lead)

#checking my predictions using race data
table(test$lead)

table(p)
```

```{r}
dim(p_class)
dim(test$lead)
x=confusionMatrix(p_class, test$lead)
```

```{r}
library(caTools)
colAUC(p,test$lead, plotROC = TRUE)

```


#Performing an ANOVA statistical test
```{r}
an <- anova(candreg)
an
```

#Clustering by race
```{r}
clusters <- kmeans(x = race[c(5:12)],centers = 6 )

# Save the cluster number in the dataset as column 'Cluster'
race$Cluster<- as.factor(clusters$cluster)

```


#Linear regression on candidates
```{r}
candreg = glm(total_votes ~ None.of.these.candidates + Alyson.Kennedy + Bradford.Lyttle+  Chris.Keniston +  Dan.Vacek+ Darrell.Castle+ Donald.Trump+   Emidio.Soltysik+  Evan.McMullin+ Frank.Atwood+  Gary.Johnson+  Gloria.La.Riva+ Hillary.Clinton+ Jerry.White+  Jill.Stein+  Jim.Hedges+   Joseph.Maldonado+  Kyle.Kopitke+  Laurence.Kotlikoff+  Lynn.Kahn+  Michael.Maturen+  Mike.Smith+ Monica.Moorehead+  Peter.Skewes+   Princess.Jacob+  Richard.Duncan+  Rocky.De.La.Fuente+   Rocky.Giordani+  Rod.Silva+  Ryan.Scott+  Scott.Copeland+   Tom.Hoefling, data = regrpesubset)

summary.lm(candreg)
```






#Exporting my pre-processed dataframes to excel
```{r}
#write.xlsx(x = pe16rwvd9, file = "pe16rwvd9.xlsx", sheetName = "TestSheet", row.names = FALSE)
#write.xlsx(x = race, file = "race.xlsx", sheetName = "TestSheet", row.names = FALSE)
#write.xlsx(x = eprace, file = "exitpollrace.xlsx", sheetName = "TestSheet", row.names = FALSE)
```

```{r}



sum(maincands[maincands$st=="AL",]$total_votes)
sum(maincands[maincands$st=="AK",]$total_votes)
sum(maincands[maincands$st=="AZ",]$total_votes)
sum(maincands[maincands$st=="CA",]$total_votes)
sum(maincands[maincands$st=="CO",]$total_votes)
sum(maincands[maincands$st=="CT",]$total_votes)
sum(maincands[maincands$st=="DE",]$total_votes)
sum(maincands[maincands$st=="FL",]$total_votes)
sum(maincands[maincands$st=="GA",]$total_votes)
sum(maincands[maincands$st=="HI",]$total_votes)
sum(maincands[maincands$st=="ID",]$total_votes)
sum(maincands[maincands$st=="IL",]$total_votes)
sum(maincands[maincands$st=="IN",]$total_votes)
sum(maincands[maincands$st=="IA",]$total_votes)
sum(maincands[maincands$st=="KS",]$total_votes)
sum(maincands[maincands$st=="KY",]$total_votes)
sum(maincands[maincands$st=="LA",]$total_votes)
sum(maincands[maincands$st=="ME",]$total_votes)
sum(maincands[maincands$st=="MD",]$total_votes)
sum(maincands[maincands$st=="MA",]$total_votes)
sum(maincands[maincands$st=="MI",]$total_votes)
sum(maincands[maincands$st=="MN",]$total_votes)
sum(maincands[maincands$st=="MS",]$total_votes)
sum(maincands[maincands$st=="MO",]$total_votes)
sum(maincands[maincands$st=="MT",]$total_votes)
sum(maincands[maincands$st=="NE",]$total_votes)
sum(maincands[maincands$st=="NV",]$total_votes)
sum(maincands[maincands$st=="NH",]$total_votes)
sum(maincands[maincands$st=="NJ",]$total_votes)
sum(maincands[maincands$st=="NM",]$total_votes)
sum(maincands[maincands$st=="NY",]$total_votes)
sum(maincands[maincands$st=="NC",]$total_votes)
sum(maincands[maincands$st=="ND",]$total_votes)
sum(maincands[maincands$st=="OH",]$total_votes)
sum(maincands[maincands$st=="OK",]$total_votes)
sum(maincands[maincands$st=="OR",]$total_votes)
sum(maincands[maincands$st=="PA",]$total_votes)
sum(maincands[maincands$st=="RI",]$total_votes)
sum(maincands[maincands$st=="SC",]$total_votes)
sum(maincands[maincands$st=="SD",]$total_votes)
sum(maincands[maincands$st=="TN",]$total_votes)
sum(maincands[maincands$st=="TX",]$total_votes)
sum(maincands[maincands$st=="UT",]$total_votes)
sum(maincands[maincands$st=="VT",]$total_votes)
sum(maincands[maincands$st=="VA",]$total_votes)
sum(maincands[maincands$st=="WA",]$total_votes)
sum(maincands[maincands$st=="WV",]$total_votes)
sum(maincands[maincands$st=="WI",]$total_votes)
sum(maincands[maincands$st=="WY",]$total_votes)

```


